An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes

### 2016-05-02 cal_demo/ms.tex
Starting major revisions in cal sec 2.1

The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, 
$E(\hat{\mathbf{s}},t)$, through a Fourier transform.
\begin{equation}
\Er{}(\mathbf{r},t) = \int E(\hat{\mathbf{s}},t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
\end{equation}
Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is 
the observer's location (measured in wavelengths relative to an arbitrary origin). We will 
encounter several quantities which we attempt to estimate. We distinguish the ``true'' values 
with a superscript $T$, while the estimated quantities are denoted with a prime. We define the 
true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric 
field on the ground.
\begin{equation}
\Er{a}^T(t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},t) \, \dif^2 \mathbf{r}
\end{equation}
The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.

>>>>

%%% New text
We begin by considering the electric field vector incident on the ground as a function of time, 
$\mathbf{E}(\mathbf{r},t)$. The position vector, $\mathbf{r}$, is measured from some arbitrary 
origin. Following \citealt{cla99} we make a few assumptions in order to clarify notation and 
simplify the problem. First we note that the time ordered electric field can be expressed as a 
Fourier series with coefficients corresponding to frequencies $f$. This allows us to analyze the 
quasi-monochromatic components of the electric field, $\mathbf{E}(\mathbf{r},f,t)$, with the 
understanding that the full field is composed of all components and their corresponding Fourier 
functions. We retain the time dependence because in practice each coefficient will only be 
measured over some sampling period.

It is convenient to simplify notation by expressing the position vector in units of wavelengths, 
$\mathbf{r}\rightarrow \mathbf{r}/\lambda$. We will also assume all antennas lie in a plane, 
reducing the problem to two dimensions. While the MOFF algorithm in principle can account for 
non-coplanar arrays through the W-projection algorithm \citep{cor08}, the EPIC code does not 
currently support this so we leave it to future work.

Next we treat the electric field as a scalar quantity, ignoring polarization effects. This allows us 
to express the electric field as a scalar, $E(\mathbf{r},f,t)$, and treat subsequent field 
propagation and measurements as scalar rather than tensor multiplications. A full polarization 
treatment is required to construct polarized sky images, however the calibration will only 
require a sky model expressed in the instrumental polarization basis. We can therefore treat 
each polarization independently.

Finally we assume the sources of the electric field are far away, allowing us to describe the 
electric fields propagating from sky sources as originating from an imaginary celestial sphere 
located a large distance from the ground. We express this intermediary electric field as an 
angular distribution, $\mathcal{E}(\hat{\mathbf{s}},f,t)$, where $\hat{\mathbf{s}}$ is the unit 
vector in the direction of a patch of the celestial sphere. We further assume the space between 
the celestial sphere and the ground is empty, allowing us to express the electric field incident 
on the ground as the integral of the propagated fields from the sphere.
\begin{equation}
E(\mathbf{r},f,t) = \int \mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega
\end{equation}  
By making the latter assumptions we have left treatment of the ionosphere and terrestrial 
emission such as local interference or thermal radiation from the ground for future work.

Next we turn to the measurement performed by an antenna with far-field radiation pattern 
$\mathcal{W}_a(\hat{\mathbf{s}})$, where the subscript $a$ denotes a specific antenna. We 
adopt the common convention of normalizing the radiation pattern to have response of 1 at the 
center of the main lobe \citep[e.g.][]{nap99}. Consequently, the translation between the terminal 
voltages to electric field measurements has been absorbed in our definition of the radiation 
pattern. The ideal antenna electric field measurement can be expressed as 
\begin{equation}
E^I_a(f,t) = \int \mathcal{W}_a(\hat{\mathbf{s}})\mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega. 
\end{equation}
Equivalently we can express the effect of the antenna radiation pattern in the plane of the 
aperture.
\begin{equation}
E^I_a(f,t) = \int_{\mathrm{aperture}}W_a(\mathbf{r}-\mathbf{r}_a)E_a(\mathbf{r},t)\,d^2\mathbf{r}
\end{equation}
Here we used the Fourier relationship between the antenna's far-field radiation pattern and the 
aperture distribution, $W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) \, d
\Omega$. Although any antenna extends over a region of space, we include a nominal antenna 
position, $\mathbf{r}_a$, for multiple reasons. In the case where all antennas of an array are 
identical, a global aperture distribution can be adopted and the translation of the antennas 
distinguishes them. However, even if all antennas have unique aperture distributions, it can be 
convenient to retain the relative positions of antennas in the analysis. This allows for easy 
comparison with visibility-based analyses where the baseline vectors are used to describe 
sampling of the $uv$ plane. From a practical point of view, measurements of the antenna 
radiation pattern may not be performed in-situ, so it is convenient to describe the antenna 
aperture distribution relative to a point fixed within the aperture, along with a translational 
position.

%The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, 
%$E(\hat{\mathbf{s}},t)$, through a Fourier transform.
%\begin{equation}
%\Er{}(\mathbf{r},t) = \int E(\hat{\mathbf{s}},t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
%\end{equation}
%Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is 
%the observer's location (measured in wavelengths relative to an arbitrary origin). We will 
%encounter several quantities which we attempt to estimate. We distinguish the ``true'' values 
%with a superscript $T$, while the estimated quantities are denoted with a prime. We define the 
%true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric 
%field on the ground.
%\begin{equation}
%\Er{a}^T(t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},t) \, \dif^2 \mathbf{r}
%\end{equation}
%The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
